Redshift Evolution of the Type Ia Supernova Population Stretch Distribution

### 2020-01-30 Nicolas.tex
Small changes; notations

\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
as a function of redshift, given our aforementioned model of the evolution of
the fraction of young and old SNe~Ia with redshift, we need to model the SN
stretch distribution for each age subsample. Given the structure of the
stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
model as follows: for the young population the underlying stretch distribution
is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
same as the young population one. \nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1\left(z\right)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1\left(z \right) =
    }&\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
\end{align}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}}\right),
\end{equation}

\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $p_y{}^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

@@

\nn{When using SNf data, we choose $p_y{}^i$ to be their given
$p_y(\mathrm{LsSFR}^i)$, computed from their LsSFR data (see
Fig.~\ref{fig:stretchlssfr}). Thus, we would have $\mathcal{P}$ from
eq.~\ref{eq:likelihood} to be}

\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
    \nn{p_y(\mathrm{LsSFR}^i)} \times &\mathcal{N}\left(\mu_1,
    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
    \nonumber\\
    (1-\nn{p_y(\mathrm{LsSFR}^i)}) \times \nn{\bigg[} a \times &
        \mathcal{N}\left(\mu_1,
        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
        \nonumber\\
     (1-a) \times &\mathcal{N}\left(\mu_2,
 \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
\end{align}

@@

Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
by fitting the free parameters of the model ($\theta\equiv{\mu_1, \mu_2,
\sigma_1, \sigma_2, a}$) assuming the evolution of the fraction of young SNe~Ia
as a function of redshift given by $\delta(z)$ and given, this time, $x_1^{i}$,
$\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch, error on the stretch and
redshift of any given SN $i$. In practice, it means that we are still minimizing
for the same eq.~\ref{eq:likelihood}, but \nn{using $p_y{}^i = \delta(z^{i})$,
which is equivalent to replacing $p_y(\mathrm{LsSFR}^i)$ with $\delta(z^i)$ in
eq.~\ref{eq:likelihoodsnf}.}

\mr{For the rest of the analysis, we will therefore fit for
eq.~\ref{eq:likelihood} using \nn{$p_y(\mathrm{LsSFR}^i)$} --~the
probability for the SN $i$ to be young~-- when available (i.e. for SNf
dataset) and $\delta(z^{i})$ --~the expected fraction of young SNe~Ia at the
SN redshift $z^{i}$~-- otherwise.}

\mr{Results on fitting the all the 569(422) SNe~Ia data (with conservative cuts)
with this model are given table~\ref{tab:modelresults} and illustrated in
Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
stretch per bin of redshift follow our redshift drift modeling. That is, when
considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
the fraction of old SNe~Ia reduces as a function of redshift. See
section~\ref{sec:results} for a more quantitative discussion.}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift.  Markers show the mean stretch measured in
        redshift bins of equal sample size indicated in light gray at the bottom
        of each redshift bins. Full and light markers are used when considering
        the nominal or the conservative sample, respectively, and are colored by
        their mean redshift.  The orange horizontal line represent the mean
        redshift of the nominal sample illustrating the expectation if the SN
        stretch distribution is not drifting. Best fits of our Base drifting
        model are shown as blue, dashed-blue and gray, when fitted on the
    nominal, the conservative or the SNf dataset, respectively; all are similar.
The light-blue illustrate the amplitude of the error of the best fit model when
considering the nominal dataset.}
    \label{fig:modelall}
\end{figure*}

@@

The ability these models to explain the data is detailed in
section~\ref{sec:results}.

@@

\begin{table*}
    \centering
    \caption{Comparison relative ability for a model to describe the data using
    AIC criterion. See the model definition section~\ref{sec:modeling}. The
drift column indicates if the model contains the age-drifting model
($\delta(z)$) or not (f, i.e. non-drifting). The BBC modeling is sample based
which somewhat contains redshift dependency information. $\mathcal{L}$ is the
pseudo-$chi^{2}$ defined eq.~\ref{eq:likelihood} and Proba is the probability
that a model is at least as good as the reference model to describe the data.
See also Fig.~\ref{fig:mod_comp}.}
    \label{tab:comp}
    \begin{tabular}{c c c  |c c c c|  c c c c}\hline\hline\\[-0.8em]
        & & & \multicolumn{4}{|c}{All SNe~Ia (569)} & \multicolumn{4}{|c}{All SNe~Ia (conservative; 422)} \\
        %\cline{2-6}
        Name & drift & Free param &
        $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba & $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba \\[0.15em]
        \hline\\[-0.8em]

@@

\section{\nn{BBC discussion}}
\label{sec:bbc}
\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
    best model, it is penalized by requiring 15 free parameters, 3 per sample.
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because current surveys span limited redshift ranges, the
    per sample approach somewhat account for unmodeled redshift drifts. We
    stress here that, as measure of modern surveys cover larger redshift ranges
    to reduce calibration systematic  uncertainties, this is becoming less true,
    notably for PS1, DES and, soon, LSST. Already with the current surveys, the
    BBC sample-based technique is excluded to be an as good representation of
    the data as our best model ($p=4\times 10^{-8}$). A more detailed discussion
    of the consequence of this results for cosmology follows in
section~\ref{sec:bbc}}

\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates our somewhat
    simple construction of our selection-bias free sample, see
    section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go from -32 to -47. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. Furthermore,
    fixing the Base model parameters to those derived using SNf data only (top
    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift level arm to probe the impact of potential redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability for a SN to be young $p(y)^i$ is
    known while other samples rely on $\delta(z^i)$ which only is the fraction
    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for $p(y)^i$.
    While in any case the Base-model is favored, the relative weight of a SNf
    sample highlight the importance of addition non-targeted nearby surveys such
    the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,graham2019} to
unambiguously confirm our result.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        distributions for the PS1 sample. The Green line(band) shows our best
        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
        approach. The dashed line shows that estimated by \cite{scolnic2018a}.
    The blue to red curves display our Base drifting stretch model estimated at
each PS SN redshift; bluer for lower redshifts.}
    \label{fig:bbc_pdf_ps1}
\end{figure}

@@

\mr{Nonetheless, to have a rough estimation of the expected amplitude of the
effect, we show Fig.~\ref{fig:magdrift} the difference of mean stretch
correction $\alpha\times\left(\langle x_1 \rangle_{\mathrm{BBC}} - \langle x_1
\rangle_{\mathrm{Base}}\right)$ as a function of redshift, using $\alpha=0.156$
from \cite{scolnic2018a}. We see in this figure that the amplitude of the
magnitude bias caused by an accurate modeling of the underlying stretch
distribution is of the order of few tens of millimag, which is the typical scale
exotic forms of dark energy could have on $\mu(z)$. In the era of modern
cosmology where we aim at probing $w_0$ at a sub-percent and $w_a$ at a few
percent \citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to
further understand the exact modeling of the lightcurve parameters and
especially their potential redshift drift; see also discussions about the impact
of stretch and color distribution modeling for cosmology in \citealt{rubin2015}
and \citealt{rubin2016}.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
    \caption{ Difference of mean stretch standardization (in mag) between the
        BBC algorithm used in \cite{scolnic2018a} and our Base drifting model.
        The blue markers represent the mean evolution in 30 regular redshift
        bins, while the small colored points show the SDSS, PS1 and SNLS
        samples. For comparison, we show in gray bands the expected magnitude
        difference between $\Lambda$CDM  and $w$CDM cosmologies ($w=-1\pm
        10,\,5\,\mathrm{and}\,1\%$) or, in orange, more exotic $w_0,w_a$CDM
        examples.  This figure illustrates that the amplitude of the potential
    magnitude bias caused by inaccurate modeling of the underlying stretch
distribution is comparable to the observational effect of exotic form of dark
energy.}
    \label{fig:magdrift}
\end{figure}

\section{Conclusion}
\label{sec:ccl}
We have presented a study of the drift of the underlyin SNe~Ia stretch
distribution as a function of redshift.  We used SNe~Ia from magnitude-limited
surveys from the pantheon data set \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
well as HST to which we added SNfactory data from \cite{rigault2018} for the
low-redshift bin. We only considered the supernovae that have been discovered
in redshift ranges of each surveys where selection effect are negligible. This
way the observed SNe~Ia stretch are random sampling of the true underlying
distribution. This corresponds the 569 SNe~Ia (422 when considering more
conservative cuts).

Following observations made in \cite{rigault2018} we introduced a redshift drift
modeling that depend on the expected fraction of young and old SNe~Ia as a
function of redshift, each age-population having its own underlying stretch
distribution. We studied various suites of distribution as well as non-drifting
modeling. We also closely studied the stretch distribution modeling made in
current version of the BCC algorithm used for Malquist bias correction by all
recent Dark energy or Hubble constant cosmological analyses.

Our conclusion are the following:
\begin{enumerate}

    \item Non-drifting models are excluded to be as good description of the data
        in comparison to our Base modeling.  This model assumes that: (1) the
        young population has a unimodal Gaussian stretch distribution while the
        old population stretch one is bimodal, one mode been the young one; (2)
        the evolution of the relative fraction of young and old SNe~Ia follows
        the prediction made in \cite{rigault2018}.

>>>>

\nn{In~\cite{rigault2018} we presented the relation between SNe stretches and
LsSFR measurements, which trace progenitor age, from the SNf sample, and is
shown in Fig.~\ref{fig:stretchlssfr}}. To model the evolution of the SN stretch
as a function of redshift, given our aforementioned model of the evolution of
the fraction of young and old SNe~Ia with redshift, we need to model the SN
stretch distribution for each age subsample. Given the structure of the
stretch--LsSFR scatter plot shown in Fig.~\ref{fig:stretchlssfr}, we define our
model as follows: for the young population the underlying stretch distribution
is modeled as $\mathcal{N}(\mu_1, \sigma_1^2)$, i.e. a normal distribution
centered on $\mu_1$ with a $\sigma_1$ width; the old population stretch
distribution is modeled as $a\times \mathcal{N}(\mu_1, \sigma_1^2) + (1-a)\times
\mathcal{N}(\mu_2, \sigma_2^2)$, i.e. a gaussian mixture where one mode is the
same as the young population one. \nn{When combining our prediction of the
evolution of young SNe~Ia as a function of redshift (eq.~\ref{eq:delta}),
and the stretch distribution of both young and old SNe~Ia, our model for the
underlying distribution of stretch $X_1\left(z\right)$ as a function of
redshift is:}
\begin{align}
    \label{eq:stretchz}
    \nn{X_1\left(z \right) =
    } &\,\delta(z)\times\mathcal{N}(\mu_1,\sigma_1^2)\,+\nonumber\\
    (1-&\,\delta(z)) \times  \left[a\times\mathcal{N}(\mu_1,\sigma_1^2) +
    (1-a)\times\mathcal{N}(\mu_2,\sigma_2^2)\right]
\end{align}

@@

\begin{equation}
    \label{eq:likelihood}
    \mathcal{L} = -2 \sum_i \ln\left( \prob{x^{i}_{1}}{ \theta;
    \mathrm{d}x^{i}_{1}, \nn{y^i}}\right),
\end{equation}

\nn{where $i$ is the index of the SN~Ia, $x^{i}_{1}$, $\mathrm{d}x^{i}_{1}$ and
    $y^i$ are the \textsc{\texttt{SALT2.4}} stretch, error of the
    \textsc{\texttt{SALT2.4}} stretch and the probability that the SN is young,
    respectively. Depending on the data available, there are two ways to fit
these 5 free parameters}.

@@

\nn{When using SNf data, we choose $y^i$ to be their given $p_y{}^i$, computed
from their LsSFR data (see Fig.~\ref{fig:stretchlssfr}). \nn{It represents the
probability to be above $\log(\mathrm{LsSFR} = -10.82)$.} Thus, we would have
$\mathcal{P}$ from eq.~\ref{eq:likelihood} to be}

\begin{align}
    \label{eq:likelihoodsnf}
    \prob{x^{i}_{1}}{\theta; \mathrm{d}x^{i}_{1}, p\nn{_y{}^i}} =
    \nn{p_y{}^i} \times &\mathcal{N}\left(\mu_1,
    \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
    \nonumber\\
        (1-\nn{p_y{}^i}) \times \nn{\bigg[} a \times &
        \mathcal{N}\left(\mu_1,
        \sqrt{\sigma_{1}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) +
        \nonumber\\
     (1-a) \times &\mathcal{N}\left(\mu_2,
 \sqrt{\sigma_2{}^{2}+\mathrm{d}x^{i}_{1}{}^{2}}\right)(x^{i}_{1}) \nn{\bigg]}
\end{align}

@@

Given eq.~\ref{eq:stretchz}, we can extend the analysis to non-SNfactory sample
by fitting the free parameters of the model ($\theta\equiv{\mu_1, \mu_2,
\sigma_1, \sigma_2, a}$) assuming the evolution of the fraction of young SNe~Ia
as a function of redshift given by $\delta(z)$ and given, this time, $x_1^{i}$,
$\mathrm{d}x^{i}_{1}$ and $z^{i}$, the stretch, error on the stretch and
redshift of any given SN $i$. In practice, it means that we are still minimizing
for the same eq.~\ref{eq:likelihood}, but \nn{using $y^i = \delta(z^i)$, which
is equivalent to replacing $p_y{}^i$ with $\delta(z^i)$ in
eq.~\ref{eq:likelihoodsnf}.}

\mr{For the rest of the analysis, we will therefore fit for
eq.~\ref{eq:likelihood} using \nn{$p_y{}^i$} --~the
probability for the SN $i$ to be young~-- when available (i.e. for SNf
dataset) and $\delta(z^{i})$ --~the expected fraction of young SNe~Ia at the
SN redshift $z^{i}$~-- otherwise.}

\mr{Results on fitting all the 569(422) SNe~Ia data (with conservative cuts)
with this model are given table~\ref{tab:modelresults} and illustrated in
Fig.~\ref{fig:modelall}. We see in this figure that the measured mean SN~Ia
stretch per bin of redshift follows our redshift drift modeling. That is, when
considering selection-bias free SN~Ia samples, SNe~Ia at higher redshift have on
average larger stretch ($0.34 \pm 0.10$ at $z\sim0.65$) than those at lower
redshift ($-0.17\pm 0.10$ at $z\sim0.05$). This is indeed what is expected if
old environments favor low SN stretch \citep[e.g.]{howell2007} \textit{and} if
the fraction of old SNe~Ia reduces as a function of redshift. See
section~\ref{sec:results} for a more quantitative discussion.}

\begin{figure*}
    \centering
    \includegraphics[width=0.7\linewidth]{Article_figures/stretchevol_all_vs_snf.pdf}
    \caption{Evolution of the mean SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        as a function of redshift. Markers show the mean stretch measured in
        redshift bins of equal sample size indicated in light gray at the bottom
        of each redshift bins. Full and light markers are used when considering
        the nominal or the conservative sample, respectively, and are colored by
        their mean redshift. The orange horizontal line represents the mean
        redshift of the nominal sample illustrating the expectation if the SN
        stretch distribution is not drifting. Best fits of our Base drifting
        model are shown as blue, dashed-blue and gray, when fitted on the
    nominal, the conservative or the SNf dataset, respectively; all are similar.
The light-blue illustrates the amplitude of the error of the best fit model when
considering the nominal dataset.}
    \label{fig:modelall}
\end{figure*}

@@

The ability of these models to explain the data is detailed in
section~\ref{sec:results}.

@@

\begin{table*}
    \centering
    \caption{Comparison of the relative ability for a model to describe the data
        using AIC criterion. See the model definition
        section~\ref{sec:modeling}. The drift column indicates if the model
        contains the age-drifting model ($\delta(z)$) or not (f, i.e.
        non-drifting). The BBC modeling is sample based which somewhat contains
        redshift dependency information. $\mathcal{L}$ is the pseudo-$\chi^{2}$
        defined eq.~\ref{eq:likelihood} and Proba is the probability that a
        model is at least as good as the reference model to describe the data.
    See also Fig.~\ref{fig:mod_comp}.}
    \label{tab:comp}
    \begin{tabular}{c c c  |c c c c|  c c c c}\hline\hline\\[-0.8em]
        & & & \multicolumn{4}{|c}{All SNe~Ia (569)} & \multicolumn{4}{|c}{All SNe~Ia (conservative; 422)} \\
        %\cline{2-6}
        Name & drift & Free param &
        $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba & $\mathcal{L}$ & $\mathrm{AIC}$ & $\Delta \mathrm{AIC}$ & Proba \\[0.15em]
        \hline\\[-0.8em]

@@

\section{\nn{BBC discussion}}
\label{sec:bbc}
\mr{Surprisingly, the BBC technique, which assumes an individual asymmetric
    distribution per sample \citep{scolnic2016,kessler2017}, is one of the worst
    modelisation. While its pseudo-$\chi^{2}$ $\mathcal{L}$ is the best of all
    the non-drift model, but still $\Delta \mathcal{L}=-11.5$ worse than our
    best model, it is penalized by requiring 15 free parameters, 3 per sample.
    \citet[][section~2]{scolnic2016} and \citet[][section~5.4]{scolnic2018a}
    highlighted that, because current surveys span limited redshift ranges, the
    per sample approach somewhat accounts for unmodeled redshift drifts. We
    stress here that, as measurements of modern surveys cover larger redshift
    ranges to reduce calibration systematic  uncertainties, this is becoming
    less true, notably for PS1, DES and, soon, LSST. Already with the current
    surveys, the BBC sample-based technique is excluded to be an as good
    representation of the data as our best model ($p=4\times 10^{-8}$). A more
    detailed discussion of the consequence of this results for cosmology follows
in section~\ref{sec:bbc}}

\mr{We report in table~\ref{tab:bbc} the samples' $\sigma^-$, $\sigma^+$ and
    $\mu^0$ that we find in close agreement with \cite{scolnic2016} for SNLS and
    SDSS and with \cite{scolnic2018a} for PS1. This validates \nn{the} somewhat
    simple construction of our selection-bias free sample, see
    section~\ref{sec:sample}. If we were to use the \cite{scolnic2016} and
    \cite{scolnic2018a} best fit values for SNLS, SDSS and PS1, respectively,
    the $\Delta\mathrm{AIC}$ between our Base drifting model and the BBC
    modeling would go from -32 to -47. \nn{It would seem that our minimization
    of these parameters is closer to our Base model than theirs}. Furthermore,
    fixing the Base model parameters to those derived using SNf data only (top
    row table.~\ref{tab:modelresults}) and using the BBC parameters from the
    aforementioned literature for PS1, SDSS and SNLS, we find
    $\Delta\mathrm{AIC}=-31.7$. Removing the SNf data reduces this difference to
    $\Delta\mathrm{AIC}=-5$. Such a reduction is expected for two reasons.
    First, it removes the nearby SNe~Ia, and thereby significantly reduces the
    redshift level arm to probe the impact of potential redshift drifts.
    Second, the SNf data are better at discriminating the individual stretch
    distributions by age since the probability for a SN to be young $p_y{}^i$ is
    known while other samples rely on $\delta(z^i)$ which only is the fraction
    of young SNe~Ia at the SN redshift, i.e. a probabilistic proxy for
    $p_y{}^i$. While in any case the Base-model is favored, the relative weight
    of a SNf sample highlights the importance of adding non-targeted nearby
surveys, such as the Zwicky Transient Facility \cite[ZTF,][]{bellm2019,
graham2019} to unambiguously confirm our result.}

@@

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/bbc_comp_PS1.pdf}
    \caption{Underlying SN \textsc{\texttt{SALT2.4}} stretch ($x_1$)
        distributions for the PS1 sample. The Green line(band) shows our best
        fitted asymmetric Gaussian distribution (its error) mimicking the BBC
        approach. The dashed line shows the noe estimated by
    \cite{scolnic2018a}. The blue to red curves display our Base drifting
stretch model estimated at each PS SN redshift; bluer for lower redshifts.}
    \label{fig:bbc_pdf_ps1}
\end{figure}

@@

\mr{Nonetheless, to have a rough estimation of the expected amplitude of the
effect, we show Fig.~\ref{fig:magdrift} the difference of mean stretch
correction $\alpha\times\left(\langle x_1 \rangle_{\mathrm{BBC}} - \langle x_1
\rangle_{\mathrm{Base}}\right)$ as a function of redshift, using $\alpha=0.156$
from \cite{scolnic2018a}. We see in this figure that the amplitude of the
magnitude bias caused by an accurate modeling of the underlying stretch
distribution is of the order of few tens of millimag, which is the typical scale
that exotic forms of dark energy could have on $\mu(z)$. In the era of modern
cosmology where we aim at probing $w_0$ at a sub-percent and $w_a$ at a few
percent \citep[e.g.,][]{lsstpaper}, it is therefore of paramount importance to
further understand the exact modeling of the lightcurve parameters and
especially their potential redshift drift; see also discussions about the impact
of stretch and color distribution modeling for cosmology in \citealt{rubin2015}
and \citealt{rubin2016}.}

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{Article_figures/BBC_distmod_w0wa.pdf}
    \caption{Difference of mean stretch standardization (in mag) between the BBC
        algorithm used in \cite{scolnic2018a} and our Base drifting model.
        \nn{The small colored points show the SDSS, PS1 and SNLS evolution,
            while the orange markers represent their mean evolution in 30
            regular redshift bins.} For comparison, we show in gray bands the
            expected magnitude difference between $\Lambda$CDM and $w$CDM
            cosmologies ($w=-1\pm 10,\,5\,\mathrm{and}\,1\%$) or, in orange,
            more exotic $w_0,w_a$CDM examples. This figure illustrates that the
            amplitude of the potential magnitude bias caused by inaccurate
        modeling of the underlying stretch distribution is comparable to the
    observational effect of exotic form of dark energy.}
    \label{fig:magdrift}
\end{figure}

\section{Conclusion}
\label{sec:ccl}
We have presented a study of the drift of the underlying SNe~Ia stretch
distribution as a function of redshift. We used SNe~Ia from magnitude-limited
surveys from the pantheon dataset \citep[][SDSS, PS1 and SNLS]{scolnic2018a} as
well as HST to which we added SNfactory data from \cite{rigault2018} for the
low-redshift bin. We only considered the supernovae that have been discovered
in redshift ranges of each surveys where selection effect are negligible. This
way the observed SNe~Ia stretch are random sampling of the true underlying
distribution. This corresponds to 569 SNe~Ia (422 when considering more
conservative cuts).

Following observations made in \cite{rigault2018} we introduced a redshift drift
modeling that depends on the expected fraction of young and old SNe~Ia as a
function of redshift, each age-population having its own underlying stretch
distribution. We studied various suites of distribution as well as non-drifting
modeling. We also closely studied the stretch distribution modeling made in
current version of the BCC algorithm used for Malmquist bias correction by all
recent Dark energy or Hubble constant cosmological analyses.

Our conclusions are the following:
\begin{enumerate}

    \item Non-drifting models are excluded to be as good descriptions of the
        data \nn{as} our Base modeling. This model assumes that: (1) the young
        population has an unimodal Gaussian stretch distribution while the old
        population stretch one is bimodal, one mode being the young one; (2) the
        evolution of the relative fraction of young and old SNe~Ia follows the
        prediction made in \cite{rigault2018}.
